---
title: "Probability (learned from edx (Probability), note: the note part are exactly the same of edx course)"
author: "Elham Sharifin"
date: "4/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Key points: Discrete Probability
•	The probability of an event is the proportion of times the event occurs when we repeat the experiment independently under the same conditions.
Pr(A)=probability of event A
•	An event is defined as an outcome that can occur when something happens by chance.
•	We can determine probabilities related to discrete variables (picking a red bead, choosing 48 Democrats and 52 Republicans from 100 likely voters) and continuous variables (height over 6 feet).


## Key points: Monte Carlo Simulations
Monte Carlo simulations model the probability of different outcomes by repeating a random process a large enough number of times that the results are similar to what would be observed if the process were repeated forever.
The sample() function draws random outcomes from a set of options.
The replicate() function repeats lines of code a set number of times. It is used with sample() and similar functions to run Monte Carlo simulations.

### example:
following is a simple example of how we can calculate the probability of an event by Monto Carlo simulation. In this event, the calculation of exact probability was simple; however, sometimes it is not the case. 
```{r}
beads <- rep(c("red","blue"),times=c(2,3)) #creates a run with 2 red and 3 blue
beads
sample(beads,1) #it shows 1 random sample of beads
sample(beads,5) #it shows 5 random sample of beads (2 reds, 3 blue)
sample(beads,5)
#sample(beads,6) #error, it will not replicate
sample(beads,6,replace=TRUE) #it will replicate
```
```{r}
B <- 1000 #how many time to draw one bead
events <- replicate(B,sample(beads,1)) #draw 1 beads, B times
#events  it creats 1000 random events of beads
tab <- table(events)
tab
prop.table(tab)
```

### sample function
we dont need to use replecate. because sample function has this advantage that can drew a lot of samples!
By default this selection is without replacement!
```{r}
sample(beads,5) #we will have 2 red and 3 blues
sample(beads,5)
sample(beads,5)
#sample(beads,6) #error: because it will do it without replacement
```
```{r}
events <- sample(beads,B,replace=TRUE) #it gives us the distibution
prop.table(table(events)) #it gives us the proportion
```
## Setting the Random Seed
The set.seed() function: because all the generated numbers are randomely selected, they are different from every person run to another one. However, if you want to make sure the results are exactly the same, we can use this function.
A popular way for picking the seed is: the year - month - day.
For example, we picked 1986 on December 20, 2018:  2018 − 12 − 20 = 1986.
```{r}
set.seed(1986)
?set.seed
```

if you run R 3.6, you can revert to the original seed setting behavior by adding the argument sample.kind="Rounding". For example:

```{r}
set.seed(1)
set.seed(1,sample.kind="Rounding") #will make R 3.6 generate a seed as in R 3.5
```
## Using the mean Function for Probability
example:
```{r}
beads <- rep(c("blue","red"),times=c(3,2))
beads
```
the probability of drawing a blue bead at random:
```{r}
mean(beads=="blue")
```
exaplanation: in facet it will evaluate if it is true or false tha we have blue: (true,true,false,false,false). so that mean only calculate the numbers is true: (1,1,1,0,0), the means of 1+1+1+0+0/5=0.6 and in fact, it gives the proportion of TRUE. The probabilities are directly related to the proportion of events that satisfy a requirement.

## Key points: Probability Distributions
The probability distribution for a variable describes the probability of observing each possible outcome.
For discrete categorical variables, the probability distribution is defined by the proportions for each group. such as beads!

## Key points: Independence
Conditional probabilities compute the probability that an event occurs given information about dependent events. 
For example, the probability of drawing a second king given that the first draw is a king is (if we have 4 series cards):

Pr(Card 2 is a king∣Card 1 is a king)=3/51 # we already have taken one card, so we have 51 cards!

```{r}
x <- sample(beads,5)
set.seed(1,sample.kind="Rounding")
x[2:5] #now the probability of red is one
```


If two events A and B are independent, 
Pr(A∣B)=Pr(A) # because their probabilities do not depend on each other

To determine the probability of multiple events occurring, we use the multiplication rule.

### Equations
The multiplication rule for independent events is:
Pr(A and B and C)=Pr(A)×Pr(B)×Pr(C)

The multiplication rule for dependent events considers the conditional probability of both events occurring:
Pr(A and B)=Pr(A)×Pr(B∣A)

We can expand the multiplication rule for dependent events to more than 2 events:

Pr(A and B and C)=Pr(A)×Pr(B∣A)×Pr(C∣A and B)

### exercise:
Suppose, we have 3 cyan, 5 magneta, and 7 yellow, we take 1 card:
```{r}
set.seed(1,sample.kind="Rounding")
cyan <- 3
magenta <- 5
yellow <- 7

 #peobability of choosing cyan in the first round:
p_cyan_1 <-  cyan / (cyan + magenta + yellow) #without replacement
p_cyan_1

#probability of not choosing cyan in the second round if the first one is selected (sampling without replacement)
p_not_cyan_without <- (magenta + yellow)/(cyan + magenta + yellow-1) #without replacement
p_not_cyan_without 

#probability of not choosing cyan in the second round if the first one is selected (sampling without replacement)
p_not_cyan_with <- (magenta + yellow)/(cyan + magenta + yellow) #without replacement
p_not_cyan_with

#the probability that the first draw is cyan and the second draw is not cyan
p_cyan_1*p_not_cyan_without
p_cyan_1*p_not_cyan_with
```
## Key points: Combinations and Permutations
paste() joins two strings and inserts a space in between.
expand.grid() gives the combinations of 2 vectors or lists.
permutations(n,r) from the gtools package lists the different ways that r items can be selected from a set of n options when order matters.
combinations(n,r) from the gtools package lists the different ways that r items can be selected from a set of n options when order does not matter.

### Code: Introducing paste() and expand.grid()
```{r}
# joining strings with paste
number <- "Three"
suit <- "Hearts"
paste(number, suit)

# joining vectors element-wise with paste
paste(letters[1:5], as.character(1:5))

# generating combinations of 2 vectors with expand.grid
expand.grid(pants = c("blue", "black"), shirt = c("white", "grey", "plaid"))
```
### Code: Generating a deck of cards
```{r}
suits <- c("Diamonds", "Clubs", "Hearts", "Spades")
numbers <- c("Ace", "Deuce", "Three", "Four", "Five", "Six", "Seven", "Eight", "Nine", "Ten", "Jack", "Queen", "King")
deck <- expand.grid(number = numbers, suit = suits)
deck <- paste(deck$number, deck$suit)

# probability of drawing a king
kings <- paste("King", suits)
mean(deck %in% kings)
```
### Code: Permutations and combinations
Correction: The code shown does not generate all 7 digit phone numbers because phone numbers can have repeated digits. It generates all possible 7 digit numbers without repeats.
```{r}
library(gtools)
permutations(5,2) # ways to choose 2 numbers in order from 1:5 without replacement (when we take a number, it cannot be appeared again)

all_phone_numbers <- permutations(10, 7, v = 0:9) #v=0:9 sjows the vector. for example it cannot be from 1:10
n <- nrow(all_phone_numbers)
index <- sample(n, 5) #it show 5 random number
all_phone_numbers[index,]  #it shows the phone number associated with the index

permutations(3,2)    # order matters
combinations(3,2)    # order does not matter
```
### Code: Probability of drawing a second king given that one king is drawn
```{r}
suits <- c("Diamonds", "Clubs", "Hearts", "Spades")
numbers <- c("Ace", "Deuce", "Three", "Four", "Five", "Six", "Seven", "Eight", "Nine", "Ten", "Jack", "Queen", "King")
deck <- expand.grid(number = numbers, suit = suits)
deck <- paste(deck$number, deck$suit)

# probability of drawing a king
kings <- paste("King", suits)
mean(deck %in% kings)

# Probability of drawing a second king given that one king is drawn
hands <- permutations(52,2, v = deck) #select from vectors that are card names (deck)
first_card <- hands[,1] #it shows the column 1
second_card <- hands[,2] #column 2 of hands
sum(first_card %in% kings)

sum(first_card %in% kings & second_card %in% kings) / sum(first_card %in% kings)
# it is equivalant with Pr(B|A)=Pr(A and B)/Pr(A) #conditional probability for dependent variables!!!
```
### Code: Probability of a natural 21 in blackjack natural 21: (natural 21: Ace+king,queen,jack,ten)
```{r}
suits <- c("Diamonds", "Clubs", "Hearts", "Spades")
numbers <- c("Ace", "Deuce", "Three", "Four", "Five", "Six", "Seven", "Eight", "Nine", "Ten", "Jack", "Queen", "King")
deck <- expand.grid(number = numbers, suit = suits)
deck <- paste(deck$number, deck$suit)

aces <- paste("Ace",suits) #the vector include all Aces
facecard <- c("King","Queen","Jack","Ten") 
facecard <- expand.grid(number=facecard,suit=suits)
facecard <- paste(facecard$number,facecard$suit) #the vector includes all facecard

hands <- combinations(52,2,v=deck) #all combination of picking 2 cards from 4 hands (13*4=52)

#probability of a natural 21 given that the ace is listed first in  'combinations'
mean(hands[,1] %in% aces & hands[,2] %in% facecard)

# for more caustion: probability of a natural 21 checking for both ace first and ace second (it means: first aces and second facecard or first facecard and second aces)
mean((hands[,1] %in% aces & hands[,2] %in% facecard)|(hands[,2] %in% aces & hands[,1] %in% facecard))
```
### Code: Monte Carlo simulation of natural 21 in blackjack
```{r}
# instead of finding the exact combination as previouse, we can use Monto Carlo and find the approximate probability
set.seed(1986)
suits <- c("Diamonds", "Clubs", "Hearts", "Spades")
numbers <- c("Ace", "Deuce", "Three", "Four", "Five", "Six", "Seven", "Eight", "Nine", "Ten", "Jack", "Queen", "King")
deck <- expand.grid(number = numbers, suit = suits)
deck <- paste(deck$number, deck$suit)

aces <- paste("Ace",suits) 

facecard <- c("King","Queen","Jack","Ten") 
facecard <- expand.grid(number=facecard,suit=suits)
facecard <- paste(facecard$number,facecard$suit) 

hands <- sample(deck,2)  #drawing two card without replacement
hands

# code for B=10,000 hands of blackjack
B <- 10000
results <- replicate(B,{
  hands <- sample(deck,2)
  (hands[1] %in% aces & hands[2] %in% facecard)|(hands[2] %in% aces & hands[1] %in% facecard)
})
mean(results)
```

## Key points: The Birthday Problem
duplicated() takes a vector and returns a vector of the same length with TRUE for any elements that have appeared previously in that vector.
We can compute the probability of shared birthdays in a group of people by modeling birthdays as random draws from the numbers 1 through 365. We can then use this sampling model of birthdays to run a Monte Carlo simulation to estimate the probability of shared birthdays.

### Code: The birthday problem
what is the chance that at least two people in a classromm of 50 peopel has the same birthday?

```{r}
# how many duplicate birthday in a group of 50
n <- 50
bdays <- sample(1:365,n,replace = TRUE)
any(duplicated(bdays))  #it is only for one sample

# Monte Carlo simulation with B=10000 replicates
# for finding the probability, we have to run it for many times and have a lot of sampels
B <- 10000
results <- replicate(B,{   # returns vector of B logical values
  bdays <- sample(1:365,n,replace=TRUE)
  any(duplicated(bdays))
})
mean(results)  #calculate the proportion or probability of duplicated birthdays
```

## Key points: sapply
Some functions automatically apply element-wise to vectors, such as sqrt() and *.
However, other functions do not operate element-wise by default. This includes functions we define ourselves.
The function sapply(x, f) allows any other function f to be applied element-wise to the vector x.

The probability of an event happening is 1 minus the probability of that event not happening:
Pr(event)=1−Pr(no event)
We can compute the probability of shared birthdays mathematically:
 
Pr(shared birthdays)=1−Pr(no shared birthdays)=1−(1×364/365×364/365× ...× (365-n+1)/365)

```{r}
# function of calculation of probability of same birth days across n people
compute_bday <- function(n,B=10000){
  same_day <- replicate(B,{
    bday <- sample(1:365,n,replace = TRUE)
    any(duplicated(bday))
  })
  mean(same_day)
}

n <- seq(1,60) 
```
### Code: Element-wise operation over vectors and sapply:
```{r}
x <- 1:10
sqrt(x) #sqrt applys on every elements of the vector x

y <- 1:10
x*y  #it multiply each elements from x on y

compute_prob(n)    # does not iterate over the vector n without sapply

x <- 1:10
sapply(x,sqrt) #this is equivalent with sqrt(x)

prob <- sapply(n,compute_bday) #element-wise application of compute_prob to n 9insteaf of reiting a loop, in R we prefer to use sapply)

plot(n,prob)  #n the size of the group
```
#### Code: Computing birthday problem probabilities with sapply
if we calculate the probability of peoples do not have the same birthday:

